---
title: "Examples using tidyquant"
output: html_notebook
---

## Step by step use of tidyquant.

This notebook shows steps in using some functionality in tidyquant.  

A good page that explains times series fitting and diagnostics is 
https://people.duke.edu/~rnau/411arim.htm

Here are some other pages with usefull information.

* https://stats.stackexchange.com/questions/108374/arima-intervention-transfer-function-how-to-visualize-the-effect
* https://stackoverflow.com/questions/25224155/transfer-function-models-arimax-in-tsa
* https://cran.r-project.org/web/packages/TSA/TSA.pdf
* https://stats.stackexchange.com/questions/17533/how-is-arma-arima-related-to-mixed-effects-modeling

Here are some useful packages.
```{r}
library(tseries)
library(lubridate)
library(tidyverse)
library(glue)
library(magrittr)
library(tidyquant)
library(shiny)
library(timetk)
```

Class notes.

* use gf_get to download finanial data
* Fit the models.
* Get predicted values and confidence interveals
* output the residual.
* Run correlations tests on the residuals.
* Note points where the residuals go outside the confidence bounds. 
* get economic data from census.
* Run fits on that data.
* Correlate times series models to other linear trend data.

## A query flow

tidyquand is an ecosystem for tidy evaluation of time series data.

First use `tq_index_options` or `tq_exchange_options` to get the possible indexes or exchanges

Options.
```{r}
index_options     <- tq_index_options()
print(index_options)
```
Exchanges.
```{r}
exchange_options  <- tq_exchange_options()
print(exchange_options)
```
Enter the index you want into `tq_index` to get all the stocks in that index.  
If you saved the list of indexes into a vector, then you can select the index you want from 
that vector.
```{r}
print(index_options[7])
```

```{r}
company_table <- tq_index(index_options[7])
print(company_table)
```
Enter the exchange you want into `tq_exchange` to get all the stocks in that exchange.
If you saved the list of exchanges in a vector, then you can select the exchange you want 
from that vector. 
```{r}
print(exchange_options[3]) 
```

```{r}
company_table_exchange <- tq_exchange(exchange_options[3])
print(company_table_exchange)
```
The function `tq_exchange` will return a list of Companies.  You can provide one or many of 
those companies to `tq_get` to retrieve a data set for that company.


Extract the data for the second company listed.
```{r}
company_table_exchange[2,]
```
Here we use `tq_get` to return a list of stock prices.  
Notice we use the output from tq_exchange as the paramter for tq_get.

```{r}
tq_get(company_table_exchange[2,],get = "stock.prices")
```
In addition to asking for *stock.prices* there are a number of other measures that can be retreived.
The list of options is returned by the function `tq_get_options`
```{r}
tq_get_options()
```
Note every options works for every possible `tq_get` entry. Here are some options that work.
Notice we can include the options as a vector.  
```{r}
i <- c(1,6)
#i <-1
print(tq_get_options()[i])
d<-tq_get(company_table_exchange[2,],get=tq_get_options()[i])
head(d)

```
It is interesting to note that both options 1 and 6 can be selected at the same time.
More than one company can also be selected at the same time.
```{r}
i <- c(1,6)
print(tq_get_options()[i])
d<-tq_get(company_table_exchange[2:3,],get=tq_get_options()[i])
head(d)

```
The result is a table of tables, or actually a tibble of tibbles.  Here the tibble of *stock.prices* for
*MMM* is selected using dplyr verbs. 
```{r}
d %>% filter(symbol == "MMM") %>% 
  select(symbol,company,last.sale.price,market.cap,ipo.year,sector,industry,stock.prices) %>%  
  unnest
```
It is also possible to use mapping (functional programming) operators such as `purrr` to run analysis on 
the results of nested tibbles.  That is not discussed here yet, but it is the main reason for the nested tibble data structure. 
That is part of the function programming approach to analysis.

Let's go back to the simple case.  We will show `tq_mutate` with `ohlc_fun` and `mutate_fun`. 
```{r}
i <-1
print(tq_get_options()[i])
```

Select the tibble for *Medical/Dental Instruments*.
```{r}
d<-tq_get(company_table_exchange[2,],get=tq_get_options()[i])
head(d)
```
The `select` and `mutate_fun` work together.  The first selects the column or variables from 
*open*, *high*, *low* and *close* and sends them to the function selected by the `mutate_fun` parameter.
In this case the `ohlc_fun` is selecting the *closing* price and sending it to the simple 
moving average function *SMA* indicated by the `mutate_fun` parameter.  The *Simple Moving Average* (*SMA*)
function requires the parameter indicating the number of periods used to calculate the average.
This *Simple Moving Average* is not and *ARIMA* moving average even though it might be possble to 
specify the same model with *ARMIA*.

The `dplr` verb *gather* is used to transpose the data from having the three columns
*close*, *SMA.15* and *SMA.50* to instead having one column called price and one column called
*type* with with either *close*, *SMA.15* or *SMA.50* depending on what value was contained by that 
row.  This is a transpose from a *wide* data format to a *long* data format.  
With a wide format, the *close*, *SMA.15* and *SMA.50* values are in two separate columns.
Witht the long format, the *close*, *SMA.15* and *SMA.50* values are in separte rows.  
So, with the wide format, values are differentiated by their columns, but in the long 
format, values are diferentiated by their rows and field dedicated to tagging the value.
Each value is tagged by the column  type to indicate if it is *close*, *SMA.15* or *SMA.50*.
```{r}

print(two_dates)
dl <-  d %>% 
  tq_mutate(select = close, mutate_fun = SMA, n=15) %>% 
  rename(SMA.15 = SMA) %>% 
  tq_mutate(select = close, mutate_fun = SMA, n=50) %>% 
  rename(SMA.50 = SMA) %>% 
  select(date,close,SMA.15,SMA.50)  %>% 
  gather(key = type, value=price, c("close","SMA.15","SMA.50"))
head(dl)

```
```{r}
# Select two dates used to display a piece of the final tibble.
two_dates <- dl %>% 
  select(date,type,price) %>% 
  filter(type == "SMA.50") %>% 
  na.omit() %>% 
  slice(1:2) %>% 
  select(date)
dl %>% inner_join(two_dates) %>% 
  group_by(type) %>% 
  sample_n(2)   %>% 
  ungroup()
```

Now ggplot is used to plot the results.  The function `scale_colour` is used to specify the 
colors used.  This is good idea because colors are very important for making the
plot readable and the designer needs to have control over this aesthetic.  
One goal in color selection is to make the color distinct even for people who are 
color blind.

Looking at this plot, in my opinion, it is clear that people who have investments need to 
look at the 50 day moving average, not the day to day close to see how their stock is
 performing.  All the historical spikes are completely irrelevent.  If you react to a movement 
 you did not predict, you are just giving your money to someone who did.  
```{r}
my_palette <- c("black", "blue", "red")
dl %>% 
  na.omit() %>% 
  ggplot() +
  aes(x=date,y=price,col = type) +
  geom_line() +
  scale_colour_manual(values = my_palette)
```






Lets look at a closup of this series.  In financial analysis, there is something called
*Technical Analysis*  This includes rules on what to do if the 50 day moving averate instecects 
with the  15 day moving average. Those intersections are clearly visible in the chart
below.
```{r}
my_palette <- c("black", "blue", "red")
dl %>% 
  na.omit() %>% 
  filter(between(date,left = lubridate::as_date('20160101'), right=lubridate::as_date('20180101'))) %>% 
  ggplot() +
  aes(x=date,y=price,col = type) +
  geom_line() +
  scale_colour_manual(values = my_palette)
```
Notice that we used  `tq_mutate(select = close, mutate_fun = SMA, n=15)`
`tq_mutate` has a small list if fuctions that can be used inside of tq_mutate.  
It does not accept any arbitrary function Here is a list of the possible functions.
Unfortunatedly I don't see and `arima`, `acf` or `pacf` function.  More research on this
is needed. There are no `forecast` functions.

```{r}
tq_mutate_fun_options()
```

It is possible to calculate lags using `xts` commands.  The series is converted to time series using
`tk_xts`.  This seems like a good way to make the conversion.
```{r}
dl %>% select(date,price,type) %>% 
  filter(type == "close") %>% 
  tk_xts(silent = TRUE) %>% 
  lag.xts(k = 1:5)
```
We can, with some manipulation use the data with the forcast package.

```{r}
dc <-dl %>% select(date,price,type) %>%  
  filter(type == "close")  %>% 
  select(date,price)  %>% 
  na.omit() %>% 
  tk_xts(silent = TRUE) %$%
  forecast::Acf(price)
```
Here the auto.arima function is used. 
```{r}
dc <-dl %>% select(date,price,type) %>%  
  filter(type == "close")  %>% 
  select(date,price)  %>% 
  na.omit() %>% 
  tk_xts(silent = TRUE) %$%
  forecast::auto.arima(price)
print(dc)
```
Do the residuals have a constant variance?
The variability appears to be increasing.
```{r}
plot(resid(dc))
```

An `Acf` of the residuals shows that `auto.arima` did a pretty good job.
```{r}
forecast::Acf(resid(dc))
```

How about the `Pacf`?
```{r}
forecast::Pacf(resid(dc))
```

The Augmented Dickey Fuller Test has a high type I error rate.  
The chance of rejecting the null hypothesis when it is true is high.
The alternate hypothesis is that the series is stationary, so in this
case, the Augmented Dickey-Fuller Test shows that the residuals are stationary 
and the model does not need more differencing.  
Running this test on the residuals is a great idea.

```{r}
tseries::adf.test(resid(dc))
```
With a smaller data set, the null is not rejected.  More data means more power.  
```{r}
tseries::adf.test(resid(dc)[1:50])
```