---
title: "Introduction to tidyquant"
runtime: shiny
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(tidyverse)
library(magrittr)
library(tidyquant)
```


Copied from Matt Dancho

tidyquant is a package that integrates other R package resourses for collecting and analyzing financial 
data into a tidy framework.

Three packages integrated
: xts, zoo, quantmod, TTR, PerfomanceAnalytics

## tq_index()
Financial index data can be collected using `tq_index`.  A list of the indexes available can 
be listed with `tq_index_options()`. 

```{r,include=FALSE}
index_options <- tq_index_options()
inputPanel(
  selectInput("index_options", label = "tq_index_options:",
              choices = index_options, selected = index_options[1])
)
```

```{r,include=FALSE}
renderDataTable({
  tq_index(input$index_options)
})
```


A list of the exchanges available through `tq_exchange()` can be listed using `tq_exchange_options()`

```{r}
exchange_options <- tq_exchange_options()
inputPanel(
  selectInput("exchange_options", label = "tq_exchange_options:",
              choices = exchange_options, selected = exchange_options[1])
)
actionButton("list_exchange","List Selected Exchange")
```

```{r,include=TRUE}
exchange_list <-eventReactive(input$list_exchange,{ 
  tq_exchange(input$exchange_options)  
})
exchange_options <-eventReactive(input$list_exchange,{ 
    exchange_list() %>%  
    select(symbol,company)  %$% 
    structure(as.character(symbol), names = as.character(company))
})

renderDataTable({   exchange_list()  })  
```

```{r,include=FALSE}
renderPrint({ exchange_options() })
```

```{r,include=FALSE}
renderUI({
   selectInput("exchange_index", label = "Exchange Index:",
              choices = exchange_options(), selected = exchange_options()[1]) 
})

```




