---
title: "Time Series Analysis"
subtitle: Session 1
output:
  html_document:
    df_print: paged
---

This R notebook provides code to accompany the first *Time Series* lecture.  

Attach libraries.  These packages are referenced below.
```{r}
library(devtools)
library(zoo)
library(timeSeries)
library(forecast)
library(tidyquant)
library(tidyverse)
library(plotly)
```

Reference
: See *Time Series Data* on the Sollers Site.

# Time Series Analysis
## Time Series Data.
Time series data represent observations of an outcome measured over time.  We will start by
reviewing facilities in R to get time series data and then to plot that data.  
Later we will learn how to model, predict and forcast time series data.  

We will use examples and note what technique each example demonostrates.  
## Gas prices. 
The package `tidyquant` has some useful facilities for getting data.  It tries to provide
a link between time series analysis packages and tidyverse packages.  

In addition to `tidyquant`, `tidyverse` is a fundemental set of packages.
The tidyverse is a name given to a set of packages and the concept of keeping data organized in
a table of one row per observation.  Data analysis requires keeping the data in a useful 
structure and the tidyverse standard is an attempt to make that happen.

Here the `tidyquant::tq_get` function is used to get the Natual Gas Index(NGI).
The data will be extracted and plotted. Note the class if the returned object.

```{r}
ngi <- tq_get("NGI",complete_cases = TRUE)
class(ngi)
```
Take a look at the last few observations of the data frame.
```{r}
tail(ngi)
```
Note that there is a date field as well as other measures.  The format is supposed to look like a table 
you would see in a spreadsheet program. It is `tidy`.

```{r}
class(ngi)
```

Next we can create a typical time series plot.  We will start with a static plot.
The function `plot` adjusts for the class of the object being plotted. 
It is a good start.

```{r}
plot(ngi)
```
We can do better with the plot by constructing our own.
Here is a fairly popular plotting function `ggplot2`. 

`ggplot` uses aesthetic (aes), which defines how the data is mapped to the grid.
Then geometry specifies how the data is displayed.
```{r}
p <- ngi %>%  
  ggplot() + 
  aes(x=date,y=open) +
  geom_point() + 
  labs(title = "Natural Gas Index", subtitle = "adjusted",
       x = "Date",y = "Open Price")
p
```
There is another package that will take the output of `ggplot` and add tool tips.
The package `plotly` adds some interactive abilities to the plot.  
`plotly` does a lot of other stuff and is probably better than ggplot.
`plotly` translates `ggplot` syntax because `ggplot` is so popular, but 
it also has its own plot layering syntax.

Here a ggplot object is converted to a plotly object.  Notice that you can hover over the
points and there is a tool bar.  
```{r}
ggplotly(p)
```
Here is a data set of rain in LA.  There is a data function that can be use to load data sets.
Package writers can include data using the R Data package. 
```{r}
library(TSA)
data(larain)
```

`larain` is a time series object. 
A time series object has a date index and a data column.
Even in the more advanced representations of a time series, 
there is a date index and data.  Here there is only one column of data,
in `tidyquant` there can be many columns of data.  
Keeping the date or time index associated with the data is what makes 
the time series structure unique.  This also makes it more difficult to 
work with than other `data.frame` objects.
```{r}
class(larain)
```
The index is the date part.
```{r}
index(larain)
```
The data can be extracted using `coredata`
```{r}
coredata(larain)
```
One important property of a time series is frequency. 
A time series has a frequency.  For example if the data is monthly and there is some periodicity,
then the frequency will be 12.  The larain data is yearly and the frequency is 1.
```{r}
frequency(larain)
```

There is a plot method for this data.  
```{r}
plot(larain,main="LA Rain Data")
```

The function actually called is `plot.zoo`. `plot` is a generic function that looks for the 
right method to fit the class of the data.
```{r}
plot.zoo(larain,main="LA Rain")
```
Base r has a `ts` function. `larain` is a ts object.
There is also a package called `zoo`.  Most more advanced time series objects 
extend `zoo`.  Extend means the take the methods of zoo and add to them.  
```{r}
larain_z <- zoo(larain)
class(larain_z)
```
```{r}
print(larain_z)
```
```{r}
plot(larain_z,main="LA Rain")
```
The zoo package alows you to do much more with the time series data 
```{r}
methods(class="ts")
```
```{r}
methods(class="zoo")
```
You can see that zoo adds a lot including date functions like `as.POSIXct` and the 
ability to handle multiple columns and conversion to a data frame using `as.data.frame`.

```{r}
methods(class="yearmon")
```
If you need a `data.frame` for things like ploting, you can use `fortify`.
The fortify function will separate the date index and the core data so the 
date can be plotted on the x axis and the core data on the y-axis.
```{r}
larain_f <- fortify(larain)
head(larain_f)
```
The result is a `data.frame`.
```{r}
class(larain_f)
```
```{r}
larain_f %>% 
  ggplot() +
  aes(x=x,y=y) +
  geom_point() +
   labs(title = "LA Rain")
```

There is an extension of the `zoo` package called `xts` for extensible time series.  

```{r}
larain_zoo <- zoo(larain) 
class(larain_zoo)
```
Here is what it looks like.
```{r}
head(larain_zoo)
```
There is a `as.data.frame` method in zoo used to convert to a data frame.
```{r}
as.data.frame(larain_zoo)
```
There is also a fortify method.
Note that the fortify method produces a column withe the date index whereas 
`as.data.frame` does not.
```{r}
larain_zoo_fortify <- fortify(larain_zoo)
larain_zoo_fortify
```
If is possible to modify the names of the result
```{r}
names(larain_zoo_fortify) <- c("Date","rain")
head(larain_zoo_fortify)
```
```{r}
larain_zoo_fortify %>% 
  ggplot() +
  aes(x=Date,y=rain) +
  geom_point() +  
  labs(title = "LA Rain")
```
There is a package called `forecast` that provides some modeling abilities.
```{r}
methods(class="forecast")
```

```{r}
forecast(larain_zoo)
```
```{r}
fit <- auto.arima(larain)
forecast(fit)
```

